{\fontfamily{ptm}\selectfont A Proximal Point framework of Nesterov type accelerated gradient}

### 2024-08-08 Notebook LaTEX/Momentum Accelerations/main.tex
>w<, big updates

    \subsection{Proximal descent inequality}
        The proximal descent inequality below is a crucial piece of inequality for deriving the behaviours of algorithms. 
        \begin{theorem}[Proximal Descent Inequality]\label{thm:ppm_descent_ineq}
            With $f: \RR^n \mapsto \overline \RR^n$ $\beta$-convex where $\beta \ge 0$, fix any $x \in \RR^n$, let $p = \hprox_f(x)$, then for all $y$ we have inequality 
            $$
                \left(f(p) + \frac{1}{2}\Vert x - p\Vert^2\right)
                - 
                \left(
                    f(y) + \frac{1}{2}\Vert x - y\Vert^2 
                \right)
                \le 
                - \frac{(1 + \beta)}{2}\Vert y - p\Vert^2. 
            $$
            Recall: $\hprox_f(x) = \argmin_{u}\left\lbrace f(u) + \frac{1}{2}\Vert u - x\Vert^2 \right\rbrace$. 
        \end{theorem}
        \begin{remark}
            We use this theorem to prove the convergence of the proximal point method. 
            See the proof (\cite{bauschke_convex_2017}, theorem 12.26). 
            The additional strong convexity index is a consequence of \hyperref[thm:str_cvx_implied]{theorem \ref*{thm:str_cvx_implied}}, item (v). 
        \end{remark}

@@

\section{Accelerated gradient descent and PPM}
    Recent works from Ahn and Sra \cite{ahn_understanding_2022} and Nesterov \cite{nesterov_lectures_2018} inspired content in this section.
    Ahn explored the interpretation of Nesterov acceleration via PPM. 
    They proposed the idea of ``similar triangle" for unifying all varieties of Nesterov accelerated gradient. 
    They used PPM to derive several variations of the Nesterov accelerated gradient algorithms. 
    Finally, they refurnished \hyperref[thm:lower_approx_ppm_convergence]{theorem \ref*{thm:lower_approx_ppm_convergence}} for the proof of convergence rate for the accelerated gradient.
    Their analysis results in relatively simple arguments that exhibits powerful extensions to several variants of the Nesterov accelerated gradient.

    \par\noindent
    Interestingly, the Nesterov accelerated gradient applies to PPM too; Guler \cite{guler_new_1992} did it two decades ago. 
    He uses the idea of a Nesterov acceleration sequence faithfully. 
    One recent development of the accelerated PPM is an algorithmic framework named: ``Universal Catalyst acceleration", proposed by Lin et al \cite{lin_universal_2015}. 
    It is an application of Guler's work in the context of variance-reduction stochastic gradient algorithms for machine learning.

    \par\noindent
    \textbf{We state our contributions}. 
    In \hyperref[sec:AG_varieties]{section \ref*{sec:AG_varieties}} 
    We stated different forms of accelerated gradient that appeared in the literatures and the equivalences between their forms. 
    They are abstract because the choice of stepsize parameters in the algorithms are unspecified. 
    Abstract forms related to the proximal gradient method were our own inventions inspired by discussions in the literatures, which includes 
    \hyperref[def:ag_prox_grad_ppm]{definition \ref*{def:ag_prox_grad_ppm}}, and 
    \hyperref[def:ag_prox_grad_tri_pt]{definition \ref*{def:ag_prox_grad_tri_pt}}. 
    In \hyperref[sec:AG_useful_information]{section \ref*{sec:AG_useful_information}}
    we prove and interpret some of the unproved claims made by Ahn, Sra \cite{ahn_understanding_2022} as good exercises and for the peace of the mind. 
    In
    \hyperref[sec:generic_ag_ppm_lyapunov_analysis]{section \ref*{sec:generic_ag_ppm_lyapunov_analysis}},
    we state two important theorems for the derivation of convergences rate for various type of algorithm using an upper bound of a Lyapunov function. 
    This part is inspired by works from Ahn, Sra \cite{ahn_understanding_2022}, but with the gradient mapping and proximal gradient added into the picture, making it distinct from all algorithms discussed in Ahn and Sra's paper. 
    \hyperref[sec:scenario_i]{Section \ref*{sec:scenario_i}} and 
    \hyperref[sec:scenario_ii]{\ref*{sec:scenario_ii}} derive the convergence rate and determine the parameters for the step sizes needed to achieve an optimal convergence rate. 
    In each section, we recover a specific variant of the Accelerated gradient and identify their appearances in the literature.

    \subsection{Preliminaries}
        We introduce some additional lemma that are crucial to the derivations of non-smooth accelerated gradient method. 
        \begin{definition}[The Gradient Mapping]
            \label{def:gradient_mapping}
            Let $g = f + g$ where $f$ is $L$-Lipschitz smooth and convex, $g$ is convex. 
            Define the proximal gradient operator
            $$
                \mathcal T_L(x) := \hprox_{L^{-1}g}(x - L^{-1}\nabla f(x)),
            $$
            then the gradient mapping is defined as
            $$
                \mathcal G_L(x) = L(x - \mathcal T_L(x)). 
            $$
        \end{definition}
        \begin{remark}
            The name ``gradient mapping" comes from \cite[(2.2.54)]{nesterov_lectures_2018}, however, Nesterov was referring to only the case when $g$ is an indicator function of a convex set in his writing. 
            Of course, in Amir Beck \cite[10.3.2]{beck_first-order_nodate}, it has the exact same definition for gradient mapping as the above. 
        \end{remark}

        \begin{lemma}[Gradient Mapping Approximates Subgradient]
            \label{lemma:grad_map_lemma_first}
            Continue from 
            \hyperref[def:gradient_mapping]{definition \ref*{def:gradient_mapping}}, 
            the gradient mapping satisfies
            \begin{align*}
                x^+ &= \mathcal T_L(x), 
                \\
                L(x - x^+) &\in  \nabla f(x) + \partial g(x^+) \ni \mathcal G_L(x). 
            \end{align*}
        \end{lemma}
        \begin{proof}
            \begin{align*}
                x^+ &= [I + L^{-1}\partial g]^{-1}\circ [I - L^{-1}\nabla f](x)
                \\
                [I + L^{-1}\partial g](x^+) &\ni [I - L^{-1}\nabla f](x)
                \\
                x^+ + L^{-1}\partial g(x^+) &\ni x - L^{-1}\nabla f(x)
                \\
                x^+ - x + L^{-1}\partial g(x^+) &\ni L^{-1}\nabla f(x)
                \\
                L(x^+ - x) + \partial g(x^+) &\ni - \nabla f(x)
                \\
                L(x - x^+) &\in \nabla f(x) + \partial g(x^+)
                \\
                \mathcal G_L(x) &\in \nabla f(x) + \partial g(x^+). 
            \end{align*}
        \end{proof}

        \begin{lemma}[Linearized Gradient Mapping Lower Bound]
        \label{lemma:grad_map_linearization}
            Continue from 
            \hyperref[lemma:grad_map_lemma_first]{definition \ref*{lemma:grad_map_lemma_first}}, 
            with $x^+ = \mathcal T_L(x)$, the gradient mapping satisfies the inequality for all $z$: 
            \begin{align*}
                h(z) &\ge
                h(z) - \frac{L}{2}\Vert x - x^+\Vert^2
                \ge h(x^+) + \langle \mathcal G_L(x), z - x\rangle. 
            \end{align*}
        \end{lemma}
        \begin{proof}
            Directly from the $L$-smoothness of $f$, convexity of $g, f$, we have the list of inequalities: 
            \begin{align*}
                &f(x^+) \le 
                f(x) + \langle \nabla f(x), x^+ - x\rangle
                + \frac{L}{2}\Vert x - x^+\Vert^2, 
                \\
                &f(x) + \langle \nabla f(x), z - x\rangle 
                \le f(z), 
                \\
                &g(x^+) \le 
                g(z) + \langle \partial g(x^+), x^+ - z\rangle. 
            \end{align*}
            Now, consider adding $g(x^+)$ to the first inequality from above we get 
            {\footnotesize 
            \begin{align*}
                f(x^+) + g(x^+) 
                &\le 
                f(x) + g(x^+) + \langle \nabla f(x), x^+ - x\rangle 
                + \frac{L}{2}\Vert x - x^+\Vert^2
                \\
                &\le 
                (f(z) - \langle \nabla f(x), z - x\rangle) + 
                \left(g(z) - \langle \partial g(x^+), x^+ - z\rangle\right)
                + 
                \langle \nabla f(x), x^+ - x\rangle + \frac{L}{2}\Vert x - x^+\Vert^2
                \\
                &= f(z) + g(z) + \langle \nabla f(x), x - z + x^+ - x\rangle
                + 
                \langle \partial g(x^+), x^+ - z\rangle + \frac{L}{2}\Vert x - x^+\Vert^2
                \\
                &= 
                h(z) + \langle \nabla f(x), x^+ - z\rangle + 
                \langle \partial g(x^+), x^+ - z\rangle
                + \frac{L}{2}\Vert x - x^+\Vert^2
                \\
                &= h(z) + \langle \nabla f(x) + \partial g(x^+), x^+ - z\rangle 
                + \frac{L}{2}\Vert x - x^+\Vert^2. 
            \end{align*}
            }
            By $\mathcal G_L(x) = L(x - x^+) \in \nabla f(x) + \partial g(x^+)$ from previous dicussion, we have 
            \begin{align*}
                h(x^+) &\le 
                h(z) + \langle \mathcal G_L(x), x^+ - z\rangle + \frac{L}{2}\Vert x - x^+\Vert^2
                \\
                &= h(z) - \langle L(x^+ - x), x^+ - x + x - z \rangle 
                + 
                \frac{L}{2}\Vert x - x^+\Vert^2
                \\
                &= 
                h(z) - L\Vert x^+ - x\Vert^2 
                + L \langle x^+ - x, x - z\rangle
                + \frac{L}{2}\Vert x - x^+\Vert^2
                \\
                &= h(z) + \langle \mathcal G_L(x), x - z\rangle - 
                \frac{L}{2}\Vert x - x^+\Vert^2. 
            \end{align*}
            Therefore, the inequality is justified. 
        \end{proof}
        \begin{remark}
            Observe that the linearization $h(x^+) + \langle \mathcal G_L(x), z - x\rangle$ is anchored at $x^+$, instead of $x$. 
            Geometrically, it's tilted and it "prefers" the sharp corners of a convex function, if, $x$ is close to a sharp corner. 
        \end{remark}

    \subsection{Varieties of Nesterov accelerated gradient}\label{sec:AG_varieties}
        Here, the acrynonym: ``AG" stands for accelerated gradient. 
        \subsubsection{AG Abstract Forms}
            In this section, we list different varieties of the Nesterov accelerated method. 
            \begin{definition}[Nestrov 2.2.7]\label{def:Nes2.2.7}
                Let $f$ be a $L$ Lipschitz smooth and $\mu\ge 0$ strongly convex function. 
                Choose $x_0$, $\gamma_0 > 0$, set $v_0 = x_0$, for iteration $k\ge 0$, it
                \begin{enumerate}
                    \item[1.] computes $\alpha_k \in (0, 1)$ by solving $L\alpha_k^2 = (1 - \alpha_k)\gamma_k + \alpha_k \mu$; 
                    \item[2.] sets $\gamma_{k + 1} = (1 - \alpha_k)\gamma_k + \alpha_k \mu$;
                    \item[3.] chooses $y_k = (\gamma_k + \alpha_k \mu)^{-1}(\alpha_k \gamma_k v_k + \gamma_{k + 1}x_k)$. Compute $f(y_k)$ and $\nabla f(y_k)$; 
                    \item[4.] finds $x_{k + 1}$ such that $f(x_{k + 1}) \le f(y_k) - (2L)^{-1} \Vert \nabla f(y_k)\Vert^2$; 
                    \item[5.] sets $v_{k + 1} = \gamma_{k+1}^{-1}((1 - \alpha_k)\gamma_kv_k + \alpha_k \mu y_k - \alpha_k \nabla f(y_k))$. 
                \end{enumerate}
            \end{definition}
            \begin{remark}
                This is in Nesterov's book \cite[(2.2.7)]{nesterov_lectures_2018}. 
                For more context, the sequence $\alpha_k \in (0, 1)$ such that $\sum_{k = 1}^{\infty}\alpha_k = \infty$, and recursively we have $\lambda_{k + 1} = (1 - \alpha_k)\lambda_k$. 
                the sequence $\lambda_k$ is called an Nesterov estimating sequence. 
                It is the most generic algorithm in his book about accelerated gradient method. 
                The genericity of the algorithm is provided by item 4., which is the a special case of the smooth descent lemma.

            \begin{definition}[Ahn Sra 6.24]\label{def:agg_ppm}
                With $f$ $\mu \ge 0$ strongly convex and $L$-Lipschitz smooth, the generic PPM form is formulated for strictly positive stepsizes $\tilde \eta_i,\eta_i$: 
                \begin{align*}
                    x_{t + 1} &= \argmin_{x} \left\lbrace
                        l_f(x; y_t) 
                        + 
                        \frac{\mu}{2}\Vert x - y_t\Vert^2
                        + 
                        \frac{1}{2\tilde \eta_{t + 1}} \Vert x - x_t\Vert^2
                    \right\rbrace, 
                    \\
                    y_{t + 1} &= \argmin_{x} 
                    \left\lbrace
                        l_f(x; y_t) + \frac{L}{2}\Vert x - y_t\Vert^2 + 
                        \frac{1}{2\eta_{t + 1}}\Vert x - x_{t + 1}\Vert^2
                    \right\rbrace. 
                \end{align*}
            \end{definition}
            \begin{remark}
                This algorithm is the same as algorithm (6.24) described by Ahn and Sra \cite{ahn_understanding_2022}. 
                Observe that by setting $\mu = 0, \tilde \eta_{t + 1} = \eta_{t + 1}$ this recovers algorithm (4.8) described in Ahn and Sra \cite{ahn_understanding_2022}. 
                Finally, Ahn, Sra 6.24 has the same form as Nesterov 2.2.7. 
                See 
                \hyperref[prop:Nes2.2.7_via_ahn_sra_6.24]{proposition \ref*{prop:Nes2.2.7_via_ahn_sra_6.24}} for a demonstration. 
            \end{remark}

            \begin{definition}[AG Proximal Gradient PPM Generic Form]
            \label{def:ag_prox_grad_ppm}
                Let $h=f + g$ be the sum of convex function $g$ and convex differentiable $f$ with $L$-Lipschitz gradient. 
                Define the proximal gradient and gradient mapping operator operator: 
                $$
                \begin{aligned}
                    \mathcal T_L^{g, f}(x) &:=  
                    \hprox_{L^{-1}g}\left(x - L^{-1}\nabla f(x)\right)
                    \\
                    &= 
                    \argmin_{u}
                    \left\lbrace
                        g(u) + f(x) + \langle \nabla f(x), u - x\rangle
                        + 
                        \frac{L}{2}\Vert u - x\Vert^2
                    \right\rbrace,
                    \\
                    \mathcal G_L^{g, f}(x) &= 
                    L\left(x - \mathcal T_L^{g, f}(x)\right). 
                \end{aligned}
                $$
                Omitting the superscript $f, g$ on $\mathcal T, \mathcal G$ for simplicity since it's clear in the context. 
                Define the linear lower bounding function for $f$ at $y$, for all $x$: 
                $$
                \begin{aligned}
                    l_h(x; y) &= h(\mathcal T_L y) + \langle \mathcal G_L(y), x - y \rangle \le h(x), 
                \end{aligned}
                $$
                with that we define the algorithm:
                $$
                \begin{aligned}
                    x_{t + 1} &= \argmin_{x} \left\lbrace
                        l_h(x; y_t) + \frac{1}{2\tilde \eta_{t + 1}} 
                        \Vert x - x_t\Vert^2
                    \right\rbrace,
                    \\
                    y_{t + 1}&= 
                    \argmin_{x}
                    \left\lbrace
                        l_h(x; y_t) + \frac{L}{2}\Vert x - y_t\Vert^2 + 
                        \frac{1}{2\eta_{t + 1}} \Vert x - x_{t + 1}\Vert^2
                    \right\rbrace.
                \end{aligned}
                $$
            \end{definition}
            \begin{observation}
                With $g \equiv 0$, the above definition reduces to 
                \hyperref[def:agg_ppm]{definition \ref*{def:agg_ppm}}
                where $\mu = 0$. 
            \end{observation}

            \begin{definition}[AG Proximal Gradient Generic Form]
            \label{def:ag_prox_grad_tri_pt}
                With $h = f + g$, where $g$ is convex, $f$ is convex and $L$-Lipschitz smooth. 
                Define proximal gradient and gradient mapping operator 
                $$
                \begin{aligned}
                    \mathcal T_L(x) 
                    &:= \hprox_{L^{-1}g}(x - L^{-1}\nabla f(x)), 
                    \\
                    x^+ &= \mathcal T_L(x), 
                    \\
                    \mathcal G_L(x) 
                    &:= L(x -  x^+). 
                \end{aligned}
                $$
                Then the algorithm updates $(y_t, x_{t + 1}, z_{t + 1})$ with expression: 
                $$
                \begin{aligned}
                    y_t^+ &= \mathcal T_L(y_t)
                    \\
                    y_t &= (1 + L\eta_t)^{-1}(x_t + L\eta_t z_t)
                    \\
                    x_{t + 1} &= x_t - \tilde \eta \mathcal G_L(y_t)
                    \\
                    z_{t + 1} &= y_t - L^{-1}\mathcal G_L(y_t)
                \end{aligned}
                $$
                for all $t\in \mathbb N$ where the base case has $y_0 = x_0$. 
            \end{definition}
            \begin{remark}
                Observe that $z_{t + 1} = y_t^+$. 
            \end{remark}

            \begin{definition}[Generic Form]\label{def:agg_tri}
                With $f$ be $L$-Lipschitz \\ 
                smooth and convex, choose any $y_0 = x_0=z_0$ and a non-negative sequence $\eta_t, \tilde\eta_t$, the algorithm admits form: 
                \begin{align*}
                    x_{t + 1} &= x_t - \tilde \eta_{t + 1} \nabla f(y_t) 
                    \\
                    z_{t + 1} &= y_t - L^{-1} \nabla f(y_t) 
                    \\
                    y_{t + 1} &= 
                    (1 + L\eta_{t + 1})^{-1}
                    (
                    x_{t + 1} + L\eta_{t + 1}z_{t + 1}
                    ). 
                \end{align*}
            \end{definition}
            \begin{remark}
                The parameter $\eta_t, \tilde\eta_{t + 1}$ are exactly the same as Ahn, Sra \cite[(6.24)]{ahn_understanding_2022}, but with the choice of $\mu = 0$. 
                This form is not explictly stated by Ahn and Sra but it's trivial to realize its presence regardless. 
                They are the same parameters for both algorithms and they are equivalent. 
                For a demonstration, see 
                \hyperref[prop:tri_form_via_ppm]{proposition \ref*{prop:tri_form_via_ppm}}. 
            \end{remark}

            \begin{definition}[Strongly Convex Generic Form]
                \quad \\
                With $f$ being $L$-Lipschitz smooth and $\mu > 0$ strongly convex then the Strongly convex generic form has updates of for iterates
                \begin{align*}
                    x_{t + 1} &= (1 + \mu\tilde \eta_{t + 1})^{-1}(x_t + \mu\tilde\eta_{t + 1}y_t^+), 
                    \\
                    z_{t + 1} &= y_t - L^{-1}\nabla f(y_t), 
                    \\
                    y_{t + 1} &= (1 + L\eta_{t + 1})^{-1}(x_{t + 1} + L\eta_{t + 1}z_{t +1}). 
                \end{align*}
            \end{definition}
            \begin{remark}
                This algorithm is equivalent to 
                \hyperref[def:agg_ppm]{definition \ref*{def:agg_ppm}}. 
                See \ref*{prop:tri_scvx_from_ahn_sra_6.24} for a demonstration. 
            \end{remark}

        \subsubsection{AG Concrete Examples}
            In this section we state variants of the Nesterov accelerated gradient method found in the literatures which are not generic.

            \begin{definition}[AG Ryu 12.1]
            \label{def:ag_ryu_12.1}
                State in chapter 12 in Ryu's writing \cite{ryu_large-scale_2022} is the following variant of accelerated gradient algorithm: 
                \begin{align*}
                    z_{t + 1} &= y_k + L^{-1}\nabla f(y_t), 
                    \\
                    y_{t + 1} &= z_{t + 1} + \frac{t - 1}{t + 2}\left(
                        z_{k + 1} - z_k
                    \right). 
                \end{align*}
                An equivalent form of the above is also provided by Ryu: 
                \begin{align*}
                    z_{t + 1} &= y_t + L^{-1}\nabla f(y_t), 
                    \\
                    x_{t + 1} &= x_t + (t + 1)(2L)^{-1}\nabla f(y_t), 
                    \\
                    y_{t + 1} &= \left(
                        1 - \frac{2}{t + 2} 
                    \right)z_{t + 1} + 
                    \left(
                        \frac{2}{t + 2}
                    \right)x_{t + 1}. 
                \end{align*}
            \end{definition}
            \begin{observation}
                The base case requires cares since it's not unique, there are more representations  depending on how the base case is define, 
                Observe that the second equivalent form is an example of \hyperref[def:agg_tri]{definition \ref*{def:agg_tri}}. 
                \par\noindent
                It is stated in Ryu's writing that the base case is $x_0 = y_0 = z_0$, 
                we think this is not the only options. 
                We remind the reader that there is not an obvious choice of $t\in \mathbb Z$ such that $x_t = y_t = z_t$ where the above algorithm remains consistent for all $t \ge -1$. 
                The choice of base case presented in Ryu's writing is a choice and not a necessity. 
                We state that one of the base case scenario that only requires knowing $y_1$.

                \par\noindent 
                We need to observe the case where $t = 0$ to understand the base case of the algorithm. 
                At each iteration $t$, Given $(x_t, y_t)$, the above formula maps to $(x_{t + 1}, y_{t + 1}, z_{t + 1})$. 
                When $t= 0$, it has $y_1 = x_1$. 
                Therefore, knowing only either $y_1$, or $x_1$ can initiate the algorithm. 
                This is an alternative scenario of the base case. 
                If, we wish to write $y_0 = x_0$ at $t = 0$ for the base case instead, then it creates the following algorithm 
                \begin{align*}
                    y_{t} &= \left(
                    1 - \frac{2}{t + 2} 
                    \right)z_{t} + 
                    \left(
                        \frac{2}{t + 2}
                    \right)x_{t}, 
                    \\
                    z_{t + 1} &= y_t + L^{-1}\nabla f(y_t), 
                    \\
                    x_{t + 1} &= x_t + (t + 1)(2L)^{-1}\nabla f(y_t). 
                \end{align*}
                It has a different representations due to a specific choice of the base case. 
                \par\noindent
                Finally, by the observation that given $(z_t, y_t)$ for any $t\ge -1$, we can solve for $x_t$ \newline via $y_t = (1 - 2/(1 + t))z_t + 1/(t + 1)x_t$, obtaining $(x_t, y_t, z_t)$, hence the algorithm can be reduced to a representations with only $(z_t, y_t)$ and an updates using only $(z_t, y_t)$ to $(z_{t + 1}, y_{t + 1})$. 
                That is the momentum form presented above. 
            \end{observation}
            \begin{remark}
                We advise the reader to read the Biliography notes by Ryu \cite[chapter 12]{ryu_large-scale_2022} of his book. 
                It goes over the full context and history for this particular variant of gradient acceleration method. 
                At the time of composing this notes, the writer is still reading on the topic of accelerated gradient.     
            \end{remark}

            \begin{definition}[Proximal Gradient variant of Ryu 12.1]
            \label{def:ag_tri_pt_form_E}
                The method appeared in Ryu's text is a special case of 
                \hyperref[def:ag_prox_grad_tri_pt]{definition \ref*{def:ag_prox_grad_tri_pt}}
                with the condition that $\tilde \eta_t = \eta_t$ for all $t \in \mathbb N$ and $g \equiv 0$ for the non-smooth part of the objective. 
                As a consequence, the algorithm has the following updates for iterates $(y_t, x_{t + 1}, z_{t + 1})$: 
                \begin{align*}
                    y_t &= (1 + L\eta_t)^{-1}(x_t + L\eta_t z_t)
                    \\
                    x_{t + 1} &= x_t - \eta_t \mathcal G_L(y_t)
                    \\
                    z_{t + 1} &= y_t - L^{-1}\mathcal G_L(y_t).
                \end{align*}
            \end{definition}

            \begin{definition}[Similar Triangle Form I]
            \label{def:ag_form_similar_tria_I}
                The similar triangle form is equivalent to the classic nesterov accelerated gradient proposed back in 1983. 
                It updates iterates $(y_t, x_{t + 1}, z_{t + 1})$ using 
                \begin{align*}
                    z_{t + 1} &= y_t - L^{-1} \mathcal G_L(y_t) 
                    \\
                    x_{t + 1} &= z_{t + 1} + L\eta_t (z_{t + 1} - z_t)
                    \\
                    y_{t + 1} &= 
                    (1 + L\eta_{t + 1})^{-1}
                    (x_{t + 1} + L\eta_{t + 1}z_{t + 1}). 
                \end{align*}
            \end{definition}
            \begin{observation}
                There are a lot of observations about this form to unpack. 
                Let's list the following observations
                \begin{enumerate}
                    \item Observe that $y_t = (1 + L\eta_t)^{-1}(x_t + L\eta_t z_t)$ is a convex combination between $x_t, z_t$. Hence $z_t, y_t, x_t$ are three collinear points. Additionally, we have $y_t - x_t = L\eta_t (z_t - y_t)$, hence $\Vert y_t - x_t\Vert/\Vert z_t - y_t\Vert = L\eta_t$. 
                    \item By the algorithm, we have $z_{t + 1} - y_t = - L^{-1} \mathcal G_L(y_t)$, and $x_{t + 1} - x_t = - \tilde \eta_{t + 1} \mathcal G_L(y_t)$, hence vector $z_{t + 1} - y_t$ parallels to $x_{t + 1} - x_t$. 
                    \item Finally, with update $x_{t +1} - z_{t + 1} = L\eta_t (z_{t +1} - z_t)$, we have three colinear points $(z_t, z_{t + 1}, x_t)$ with $\Vert z_{t + 1} - z_t\Vert/\Vert z_{t +1 } - z_t\Vert = L\eta_t$. 
                \end{enumerate}
                From the above results, we can conclude that triangle $(y_t, z_t, z_{t + 1})$ is similar to $(x_t, z_t, x_{t + 1})$ because they share colinear points $z_t, y_t, x_t$ and $z_t, z_{t + 1}, x_{t + 1}$, and their sides $z_{t + 1} - y_t$ parallels to $x_{t + 1} - x_t$. 
            \end{observation}
            \begin{remark}
                The above algorithm is a special case of 
                \hyperref[def:ag_prox_grad_tri_pt]{definition \ref*{def:ag_prox_grad_tri_pt}}, 
                see 
                \hyperref[sec:scenario_ii]{section \ref*{sec:scenario_ii}} for more information. 
            \end{remark}

            \begin{definition}[Similar Triangle Form II]
            \label{def:ag_form_similar_tria_II}

            \end{definition}

        \subsubsection{Useful Observations about these Generic forms}
        \label{sec:AG_useful_information}

            In this section we list some incredibaly useful information about these forms of the Nestrov type accelerated gradient algorithm. 
            \hyperref[prop:tri_form_via_ppm]{Proposition \ref*{prop:tri_form_via_ppm}}, 
            \hyperref[prop:tri_scvx_from_ahn_sra_6.24]{proposition \ref*{prop:tri_scvx_from_ahn_sra_6.24}}, 
            \hyperref[prop:Nes2.2.7_via_ahn_sra_6.24]{proposition \ref*{prop:Nes2.2.7_via_ahn_sra_6.24}}
            are results assumed as true in Ahn, Sra. 
            They are implicitly left as necessary exercises for the readers. 
            Here, we fill out the blanks with our own interpretations. 
            \hyperref[prop:derive_ag_prox_grad_tript]{Proposition \ref*{prop:derive_ag_prox_grad_tript}} 
            is our contribution where we intorduce non-smooth ness to the objective function while still retaining the PPM interpretation of accelerated gradient.

            \begin{proposition}[Tri-points generic form via Ahn Sra 6.24]
                \label{prop:tri_form_via_ppm}
                with $f$ being \\
                $L$-Lipschitz smooth, 
                let $l_f(x, \bar x) = f(\bar x) + \langle \nabla f(\bar x), x - \bar x\rangle$, we can show that the AG generic triangular form is a consequence of th AG generic PPM form. 
            \end{proposition}
                Solving the optimality on the first PPM yields: 
                    \mathbf 0 &= \nabla f(y_t) + 
                    \frac{1}{\tilde \eta_{t + 1}} (x - x_t)
                    x &= x_t - \tilde \eta_{t + 1} \nabla f(y_t).
                Therefore, $x_{t + 1} = x_t - \tilde \eta_{t + 1}\nabla f(y_y)$. 
                Similarly, for the updates of $y_{t + 1}$, we have optimality condition of 
                    \mathbf 0 &= \nabla f (y_t) + L (x - y_t) + \eta_{t + 1}^{-1} (x - x_{t + 1})
                    \mathbf 0 &= \eta_{t + 1}\nabla f (y_t) + \eta_{t + 1}L (x - y_t) + x - x_{t + 1}
                    \mathbf 0 &= 
                    \eta_{t + 1}\nabla f(y_t) -\eta_{t + 1} Ly_t + (\eta_{t + 1}L + 1)x - x_{t + 1}
                    (1 + \eta_{t + 1}L)x
                    x_{t + 1} - \eta_{t + 1}\nabla f(y_t) + \eta_{t + 1}L y_t
                    \text{define: } y_{t + 1} &:= x. 
                In the above expression, it hides a step of gradient descent, continuing it we have 
                    (1 + \eta_{t + 1}L)y_{t + 1} &= 
                    x_{t + 1}  + \eta_{t + 1}L (-L^{-1}\nabla f(y_t) + y_t)
                    \text{let: } z_{t + 1} &= y_t - L^{-1}\nabla f(y_t), \text{ so, }
                    (1 + \eta_{t + 1}L)y_{t + 1} &= 
                    x_{t + 1} + L\eta_{t + 1}z_{t + 1}. 
                Combining it yields the tree points update format 
                    x_{t + 1} &= x_t - \tilde \eta_{t + 1} \nabla f(y_t) 
                    z_{t + 1} &= y_t - L^{-1} \nabla f(y_t) 
                    y_{t + 1} &= 
                    (1 + L\eta_{t + 1})^{-1}
                    (
                    x_{t + 1} + L\eta_{t + 1}z_{t + 1}
                    ), 
                the ordering of $x_{t +1}, z_{t + 1}$ can be permuted. 
                The base case is when $t = 0$, so then $x_0 = y_0$ for the initial guess.
                It is quite obvious to us that the choice of $z_{t + 1} = y_t - L^{-1}\nabla f(y_t)$ is deliberate. 
                In fact, it's true that $z_{t + 1}$ is just a term such that it makes $y_{t +1}$ a convex combination between the vector $x_{t + 1}, z_{t + 1}$, and the choice here would be unique because of the consideration that 
                    & \quad 
                    l_f(x; y_t) + \frac{L}{2}\Vert x - y_t\Vert^2 
                    f(y_t) + \langle \nabla f(y_t), x - y_t\rangle 
                    \frac{L}{2}
                    \left\Vert 
                        x - (y_t - L^{-1}\nabla f(y_t)) - L^{-1}\nabla f(y_t)
                    \right\Vert^2
                    f(y_t) + 
                    \langle \nabla f(y_t), x - (y_t - L^{-1}\nabla f(y_t)) - L^{-1}\nabla f(y_t)\rangle 
                    \frac{L}{2}
                    \left\Vert 
                        x - (y_t - L^{-1}\nabla f(y_t)) - L^{-1}\nabla f(y_t)
                    \right\Vert^2
                    &\quad  \text{Let }z_{t + 1} 
                    = y_t - L^{-1}\nabla f(y_t)
                    &= f(y_t) + \langle \nabla f(y_t), x - z_{t + 1} - L^{-1}\nabla f(y_t)\rangle
                    \frac{L}{2}\left\Vert
                        x - z_{t + 1} - L^{-1}\nabla f(y_t)
                    \right\Vert^2
                    f(y_t) + \langle \nabla f(y_t), x - z_{t + 1}\rangle 
                    - L^{-1}\Vert \nabla f(y_t)\Vert^2
                    \frac{L}{2}\Vert x - z_{t + 1}\Vert^2 + 
                    \frac{1}{2L}\Vert \nabla f(y_t)\Vert^2 - 
                    L\langle L^{-1}\nabla f(y_t), x - z_{t + 1}\rangle
                    &= f(y_t) + (1/(2L)- L^{-1})\Vert \nabla f(y_t) \Vert^2 + 
                    \frac{L}{2}\Vert x - z_{t + 1}\Vert^2
                    &= f(y_t) - \frac{1}{2L}\Vert \nabla f(y_t)\Vert^2 + \frac{L}{2}\Vert x - z_{t + 1}\Vert^2. 
                Therefore 
                    y_{t + 1} &= \argmin_{x}\left\lbrace
                        \frac{L}{2}\Vert x - z_{t + 1}\Vert^2 + 
                        \frac{1}{2\eta_{t + 1}}\Vert x - x_{t + 1}\Vert^2
                    \right\rbrace
                    &= \argmin_{x}\left\lbrace
                        \left\Vert x - 
                            \frac{Lz_{t + 1} + \eta_{t + 1}^{-1}x_{t + 1}}{
                                L + \eta_{t + 1}^{-1}
                        \right\Vert^2
                    \right\rbrace
                    \frac{
                        L\eta_{t + 1}z_{t + 1} + 
                        x_{t + 1}
                    }{
                        1 + L \eta_{t + 1}
                    }. 

>>>>

    \subsection{Proximal descent inequality}
        The proximal descent inequality below is a crucial piece of inequality for deriving the behaviours of algorithms. 
        \begin{theorem}[Proximal Descent Inequality]\label{thm:ppm_descent_ineq}
            With $f: \RR^n \mapsto \overline \RR^n$ $\beta$strongly convex where $\beta \ge 0$, fix any $x \in \RR^n$, let $p = \hprox_f(x)$, then for all $y$ we have inequality 
            $$
                \left(f(p) + \frac{1}{2}\Vert x - p\Vert^2\right)
                - 
                \left(
                    f(y) + \frac{1}{2}\Vert x - y\Vert^2 
                \right)
                \le 
                - \frac{(1 + \beta)}{2}\Vert y - p\Vert^2. 
            $$
            Recall: $\hprox_f(x) = \argmin_{u}\left\lbrace f(u) + \frac{1}{2}\Vert u - x\Vert^2 \right\rbrace$. 
        \end{theorem}
        \begin{remark}
            We use this theorem to prove the convergence of the proximal point method. 
            See the proof (\cite{bauschke_convex_2017}, theorem 12.26). 
            The additional strong convexity index is a consequence of \hyperref[thm:str_cvx_implied]{theorem \ref*{thm:str_cvx_implied}}, item (v). 
        \end{remark}

@@

\section{Accelerated gradient descent and PPM}
    \input{sections/ag_ppm.tex}
